A Modular Framework for Developing, Deploying, and Evaluating Game-Theoretic Strategy Design Exercises

### 2018-05-07 paper.tex
Much revision and starting results

\title{
A Modular Framework for Developing, Deploying, and Evaluating Game-Theoretic Strategy Design Exercises}

@@

While the strategy design exercises assignments are popular with students\footnote{https://reg-captiva.princeton.edu/chart/index.php?terminfo=1184\&courseinfo=012095 requires CAS login}, they also demand considerable instructor resources be dedicated to building and grading software resources, which have historically been significantly rebuilt for each assignment.
Worse still, due to course staffing changes, this work has not been preserved between semesters, resulting in inefficient allocation of department resources to producing new exercises without the institutional memory or software resources previously developed.
Due to the limited tooling provided in past semesters for students to test their code, many submissions did not compile or caused runtime exceptions, requiring considerable additional instructor resources.

% the "programming assignments" aren't really about your ability to code. They are about placing the theoretical results from the course in context. They create a scenario in which mastery of theoretical results from the course content is demonstrated by making and justifying real-world decisions.

@@

\section*{Approach}
Cyril Zhang completed the TA requirement for the PhD program last fall, so at the start of this semester it was unclear who among the graduate and undergraduate course staff would be responsible for maintaining the COS 445 strategy design exercises.
Therefore, the course needed not just long-term work to improve student and instructor resources, but also short-term work to maintain and debug the same resources.
These coinciding requirements inspired me to design one extensible implementation of all the reusable components for software design exercises, and to develop the project iteratively in conjunction with its usage in the course.
This software would naturally grow throughout development as demanded by the needs of the course, and the final product will naturally lower-bound the maximum degree of automation and modularity possible.

% TODO: make it fit
The first key decision in the implementation of these programming assignments was language selection.
Many students suggested the assignments be implemented in Python,\footnote{per survey 1 results} and this would be preferred by the course staff as well.
Members of the Mechanism Design group are not expected to have working proficiency with Java, but many have enough experience with Python to support students.
Additionally, Python's conciseness and ability to load classes from user input would be useful when implementing tools which are parameterized on user-input strategy names.
Unfortunately, COS 445's prerequisite courses only provide Java experience, and not Python experience, so students reasonably expect they will only be asked to write Java code.
Data from other course surveys indicate that 90\% of students report themselves fluent in Java after the Princeton introductory sequence, compared to 30\% for Python, confirming that the student work for the assignment must be in Java.\footnote{COS 333 survey data}
Given this constraint, and lacking any other language selection pressures, I elected to implement as much of the assignment code as possible in Java, with limited usage of scripting languages such as Python, Bash, and GNU Make.
These scripting languages were used only where Java would not suffice and only where implementation details were irrelevant to student assignment comprehension.

@@

In order to increase the reusability, I elected to store this work in a University-owned private GitHub repository and documented the repository structure and contents.
This ensures that future course staff will have the physical and technical means to maintain this project after I have graduated.

\section*{Implementation}
% TODO: signposting
\subsection*{Preliminary Choices}
% TODO: make it fit
Before the first problem set of the course was assigned in this semester, Professor Weinberg and I designed a handout interface and tools to allow students to test their strategies against each other, which I then implemented.
The game simulates college admissions using stable matchings.
While I implemented this handout with the explicit goal of producing a tool to test out multiple different student strategies in the specified application environment, I followed the principles of modular software design taught in COS 217.\footnote{Kernighan and Pike chapter 4}
I factored out the code which implemented the Gale-Shapley deferred acceptance algorithm for stable matchings from the code which performed randomized trials of the stable matchings and produced performance statistics for students.

@@

However, because of conscious design decisions to factor out repeated code from assignment implementations, it accomplishes what previous implementations had not.
It will be modular, as described above, so that new extensions such as additional tools or interfaces can be written for strategy design exercises.
Throughout the semester, I can leverage my role as an undergraduate course assistant for COS 445 during the current semester in order to identify and produce such extensions.
It will be reusable, both for the same assignments in future years and for new strategy design exercises implemented by future course staff with lower overhead.

@@

\subsection*{Auto-Grader Design}
The two most important components of each exercise are the handout code for the students and the auto-grader code for the instructors.
However, in order to design the auto-grader, Professor Weinberg and I had to develop a well-specified, portable standard for grading strategy design exercises.
A key constraint in the grading of these assignments is that the incentive for the students must be to try to maximize their own strategy's performance.
Therefore, curving each student's grade based on the percentile of their strategy's performance is not acceptable.
In this scenario, a student would be as happy to hurt their peers as to improve their own strategy.
For instance, in an iterated prisoner's dilemma, if students are graded based on relative performance, then they are apathetic between both cooperating and both defecting.
So there is no incentive under any circumstance to cooperate, even in the face of potentially irrational counterparties, and every student should always defect.

@@

In summary, each assignment consists of a strategy interface, implementations of that strategy, a list of those implementations, a \texttt{runTrial} method which specifies how to evaluate those strategies, methods which run specific subsets of strategies, methods that post-process the results of those trials to provide low-latency feedback or for auto-grading, and automating these steps with a Makefile to configure, compile, and run the assignment simulation.
\footnote{TODO redraw using Powerpoint}
\footnote{TODO add formatBase.py config csv}
\footnote{TODO use this diagram when I talk about modularization earlier}
\begin{figure}[hbt]
\centering
\includegraphics[width=0.75\linewidth]{Handwritten_Diagram.jpg}
\caption{TODO: This is a gray image.}\label{fig:gray}
\end{figure}

@@

Thus far, I have illustrated the strengths of my project at performing tasks which were envisioned from the outset.
The strength of modular implementation also comes from the ability to adapt to changing demands easily in real time.
In this way, I was able to extend my framework throughout the semester in order to address my needs as a course assistant to remedy issues with the strategy design exercises.
As shown below,\footnote{results section ref here} the handout code enabled students to verify that their code compiled and ran.
However, several students submitted non-functioning strategies due to environmental differences, demonstrating the classic issue ``Works On My Machine''.

@@

As the goals of a project naturally define evaluation criteria, I have evaluated my project on the basis of each those goals: decreasing the work to build each assignment within a semester and the instructor time allocated to assignment implementation, eliminate the inter-year assignment reconstruction, improve available student resources, and in doing so improve student and instructor satisfaction.

\begin{itemize}
\item discover modularity and decrease assignment to assignment rewriting
  \begin{itemize}
  \item code rewriting measurements
  \item instructor time usage
  \end{itemize}
\item improve student resources
  \begin{itemize}
  \item list resources created
  \item polling data about student usage
  \item empirical results about student error rates
    ``Illegal submissions'' constitute student strategies which do not compile, throw exceptions, violate assertions, or otherwise behave outside of the parameters of the simulation.\footnote{Poorly performing strategies which obey the specification are legal.}
    The illegal submission rate measures the percentage of students who were unable to develop a quality strategy for reasons other than lack of mastery of the course content.
  \end{itemize}
\item create institutional memory and decrease year-to-year rewriting
  \begin{itemize}
  \item cant measure until next year, but set up for it
  \end{itemize}
\item increase student and instructor satisfaction
  \begin{itemize}
  \item polling data about student usage
  \end{itemize}
\end{itemize}

%Thus, quantitative metrics, including software failure rates and instructor time usage, and qualitative metrics, including student and instructor satisfaction, both illustrate the improvements to the course and emphasize the importance of quality tooling to quality programming-based assignments.

%What are your thoughts on how well the strategy design exercises have gone this year? Specifically, I'm interested in your satisfaction with the implementations and software, and the amount of instructor time and frustration spent. (And less intersted in comments on the course policy decisions that may have impacted how well they went, sicne that's not in the scope of my independent work.)
%** significantly** improved instructor involvement.

%I didn't actually touch any of the software directly, so I don't know how to comment on that. But I felt like the challenges mostly ran themselves. We talked briefly to set it up. We talked briefly to decide how to grade, and I didn't have to answer any technical questions for students in between

>>>>

\title{
  A Modular Framework for Developing, Deploying, and Evaluating Game-Theoretic Strategy Design Exercises
}

@@

These strategy design exercises are intended to help students place theoretical results from the course in context.
They create a scenario in which mastery of theoretical course content is demonstrated by making and justifying real-world decisions.
Professor Weinberg's intent is that no student should experience difficulty implementing their strategy, and all of the challenge should come from developing a creative, successful strategy.

While the strategy design exercises assignments are popular with students\footnote{https://reg-captiva.princeton.edu/chart/index.php?terminfo=1184\&courseinfo=012095 requires CAS login}, they also demand dedication of considerable instructor resources to building software for student handouts and strategy evaluation, which have historically been significantly rebuilt for each assignment.
Worse still, due to course staffing changes, this work has not been preserved between semesters, resulting in inefficient allocation of department resources to producing new exercises without the institutional memory or software resources previously developed.
Due to the limited tooling provided in past semesters for students to test their code, many submissions did not compile or caused runtime exceptions, requiring considerable additional instructor resources.

@@

\section*{Approach}
Cyril Zhang completed the TA requirement for the PhD program last fall, so at the start of this semester it was unclear who among the graduate and undergraduate course staff would be responsible for maintaining the COS 445 strategy design exercises.
Therefore, the course needed not just long-term work to improve student and instructor resources, but also short-term work to maintain and debug the same resources.
These coinciding requirements inspired me to design one extensible implementation of all the reusable components for software design exercises, and to develop the project iteratively in conjunction with its usage in the course.
This software would naturally grow throughout development as demanded by the needs of the course, and the final product will naturally lower-bound the maximum degree of automation and modularity possible.
While this decision to develop my software using agile methods obviates the need for an extensive design process, a few key decisions were still needed before implementation could commence.

While I was initially inclined to pursue implementation in Java, I spent considerable effort on language selection to verify this decision.
This was necessary because I would incur a considerable cost to myself and to others ere I to change programming languages partway through the semester.
I would have to rebuild all previous work, and I would inconvenience all the students of COS 445 by making them change languages and wasting their time learning the abandoned language.

Many students suggested the assignments be implemented in Python,\footnote{per survey 1 results} and this would be preferred by the course staff as well.
Members of the Mechanism Design group are not expected to have working proficiency with Java, but many have enough experience with Python to support students.
Additionally, Python's conciseness and ability to load classes from user input would be useful when implementing tools which are parameterized on user-input strategy names.
Unfortunately, COS 445's prerequisite courses only provide Java experience, and not Python experience, so students reasonably expect they will only be asked to write Java code.
Data from other course surveys indicate that 90\% of students report themselves fluent in Java after the Princeton introductory sequence, compared to 30\% for Python, confirming that the student work for the assignment must be in Java.\footnote{COS 333 survey data}
Given this constraint, and lacking any other language selection pressures, I elected to implement as much of the assignment code as possible in Java, with limited usage of scripting languages such as Python, Bash, and GNU Make.
These scripting languages were used only where Java would not suffice and only where implementation details were irrelevant to student assignment comprehension.

@@

While of lesser urgency, I have also planned ahead to ensure the reusability of my work.
I elected to store this work in a University-owned private GitHub repository.
I also documented the repository structure and contents, including a summary of the process of developing and deploying a new assignment.
This ensures that future course staff will have the physical and technical means to maintain this project after I have graduated.

\section*{Implementation}
% TODO: signposting
% TODO: make it fit
Before the first problem set of the course was assigned in this semester, Professor Weinberg and I designed a handout interface and tools to allow students to test their strategies against each other, which I then implemented.
The game simulates college admissions using stable matchings.
While I implemented this handout with the explicit goal of producing a tool to test out multiple different student strategies in the specified application environment, I followed the principles of modular software design taught in COS 217.\footnote{Kernighan and Pike chapter 4}
I factored out the code which implemented the Gale-Shapley deferred acceptance algorithm for stable matchings from the code which performed randomized trials of the stable matchings and produced performance statistics for students.

@@

However, because of conscious design decisions to factor out repeated code from assignment implementations, my implementation has accomplished what previous implementations had not.
It is modular, as described above, so that new extensions such as additional tools or interfaces can be written for strategy design exercises.
Throughout the semester, I have leveraged my role as an undergraduate course assistant for COS 445 during the current semester in order to identify and produce such extensions.
It is reusable, both for the same assignments in future years and for new strategy design exercises implemented by future course staff with lower overhead.
\footnote{TODO redraw using Powerpoint}
\footnote{TODO use this diagram when I talk about modularization earlier}
\begin{figure}[hbt]
  \centering
  \includegraphics[width=0.75\linewidth]{Handwritten_Diagram.jpg}
  \caption{TODO: This is a gray image.}\label{fig:gray}
\end{figure}

@@

\subsection*{Auto-Grader Design}
The two most important components of each exercise are the handout code for the students and the auto-grader code for the instructors.
However, in order to design the auto-grader, Professor Weinberg and I had to develop a well-specified, portable standard for grading strategy design exercises.
A key constraint in the grading of these assignments is that the students must be incentivized to maximize their own strategy's performance, and not incentivized to either help or hurt their peers except when doing so helps themselves.\footnote{Of course, students are encouraged to help their peers out of their own compassion, and doing so contributes to their course participation grade, but they are not rewarded for doing so in their assignment.}
Therefore, curving each student's grade based on the percentile of their strategy's performance is not acceptable.
In this scenario, a student would be as happy to hurt their peers as to improve their own strategy.
For instance, in an iterated prisoner's dilemma, if students are graded based on relative performance, then they are apathetic between both cooperating and both defecting.
So there is no incentive under any circumstance to cooperate, even in the face of potentially irrational counterparties, and every student should always defect.

@@

In summary, each assignment consists of a strategy interface, implementations of that strategy, a list of those implementations, a \texttt{runTrial} method which specifies how to evaluate those strategies, methods which run specific subsets of strategies, methods that post-process the results of those trials to provide low-latency feedback or for auto-grading, and automating these steps with a Makefile to configure, compile, and run the assignment simulation.\footnote{add ref to above diagram here}
\footnote{TODO add formatBase.py config csv}

@@

Thus far, I have illustrated the strengths of my project at performing tasks which were envisioned from the outset.
The strength of modular implementation also comes from the ability to adapt to changing demands easily in real time.
In this way, I was able to extend my framework throughout the semester in order to address my needs as a course assistant to remedy issues with the strategy design exercises.
While the handout code enabled students to verify that their code compiled and ran on their own personal computers, several students submitted non-functioning strategies due to environmental differences, demonstrating the classic issue ``Works On My Machine''.

@@

While I have been unable to experimentally determine whether my project will be reused, I have evaluated my successful achievement of each other goal.
I have decreased the work to build each assignment within a semester and the instructor time allocated to assignment implementation, eliminated the inter-year assignment reconstruction, and improved available student resources
These achievements have increased student and instructor satisfaction with the strategy design exercises.

My primary goal, which unifies my work and sets it apart from past work on this subject, has been to discover modularity and decrease assignment to assignment rewriting.
\begin{itemize}
\item code rewriting measurements between assignments
\item instructor time usage
\item extensibility:
\end{itemize}

I have also succeeded at leveraging the extensibility of my project to improve student resources.

\begin{itemize}
\item list resources created
\item polling data about student usage
\item empirical results about student error rates
  ``Illegal submissions'' constitute student strategies which do not compile, throw exceptions, violate assertions, or otherwise behave outside of the parameters of the simulation.\footnote{Poorly performing strategies which obey the specification are legal.}
  The illegal submission rate measures the percentage of students who were unable to develop a quality strategy for reasons other than lack of mastery of the course content.
\end{itemize}

Through this work, I have improved student and instructor experiences.
In addition to the data gathered above, I verified this achievement by surveying the current students of COS 445 under the supervision of Professor Weinberg.
\begin{itemize}
\item polling data about student usage
\end{itemize}

% Thus, quantitative metrics, including software failure rates and instructor time usage, and qualitative metrics, including student and instructor satisfaction, both illustrate the improvements to the course and emphasize the importance of quality tooling to quality programming-based assignments.

% What are your thoughts on how well the strategy design exercises have gone this year? Specifically, I'm interested in your satisfaction with the implementations and software, and the amount of instructor time and frustration spent. (And less intersted in comments on the course policy decisions that may have impacted how well they went, sicne that's not in the scope of my independent work.)

% ** significantly** improved instructor involvement.
% I didn't actually touch any of the software directly, so I don't know how to comment on that. But I felt like the challenges mostly ran themselves. We talked briefly to set it up. We talked briefly to decide how to grade, and I didn't have to answer any technical questions for students in between
